---
title: "A novel mass spectrometry-based method for the identification of subtype specific amyloidogenic proteins"
output:

editor_options:
  chunk_output_type: inline
---

Contained in this document is the R scripts necessary to reproduce the results from Palstroem et al. "A novel mass spectrometry-based method for the identification of subtype specific amyloidogenic proteins". This document was created under R 4.0.0 and RStudio 1.2.5001.


-----------------------------------------
               CONTENTS
-----------------------------------------
 
* Preparation steps
  + Load packages
  + Import data
  + Remove NAs, and whatever

* Creating training and test dataset

* Feature selection - Boruta
  + Step 1 - Differentiating between Congo Red positive and negative samples
  + Step 2 - Recognizition of subtypes

* Establish optimal model parameters (Tuning):
  + Step 1
  + Step 2

* Model building
  + Step 1 
    + Extra proteins identified by Boruta
  + Step 2

* Test set
  + Step 1 
  + Step 2

* Validation set
  + Step 1
  + Step 2

-----------------------------------------
-----------------------------------------

#Preparation steps

###Install required packages
```{r Install required packages, eval=FALSE, include=FALSE, error=TRUE}
install.packages(c("readxl","dplyr","Boruta","caret","e1071"),repos="https://cloud.r-project.org")
```
###Load required packages
```{r Load packages, error=TRUE}
library(readxl) #Required for import of Excel data
library(dplyr) #Required for some of the data wrangling
library(Boruta) #Required for Feature Selection
library(caret) #Required for Data Partitioning
library(e1071) #Required for building the Support Vector Machine model
```


###Import the provided main dataset
```{r Main dataset import,error=TRUE}
Main_Dataset<-read_excel(file.choose(),sheet = "Main_Dataset") #Select the provided .xlsx document from the folder to which it is located
```

###Import the validation dataset
```{r Validation dataset import, error=TRUE}
Validation_Dataset<-read_excel(file.choose(),sheet = "Validation_Dataset") #Select the provided .xlsx document from the folder to which it is located
```

###Define the columns "Status" and "Diagnosis" as factors in the Main Dataset
```{r,error=TRUE}
#Main dataset
Main_Dataset$Status<-as.factor(Main_Dataset$Status)
Main_Dataset$Diagnosis<-as.factor(Main_Dataset$Diagnosis)
```

... And the Validation Dataset
```{r error=TRUE}
#Validation data
Validation_Dataset$Diagnosis<-as.factor(Validation_Dataset$Diagnosis)
Validation_Dataset$Status<-as.factor(Validation_Dataset$Status)
```

#Training and test dataset

###Step 1

####Partition the main dataset into a training dataset and a test dataset according to the "Status" column. 

*The data is randomly split with 70% allocated to the Training dataset and the remaining 30% is allocated to the Test Dataset*

```{r,error=TRUE}
set.seed(1) #Setting the seed for the pseudorandom number generator
index<- createDataPartition(Main_Dataset$Status, p = 0.7, list = FALSE) #caret package must be loaded
Training_Data <- Main_Dataset[index,]
Test_Data <- Main_Dataset[-index,]
```

###Step 2

####Remove Congo-Red negative samples from the dataset to build model for recognition of subtypes:
```{r,error=TRUE}
Data_only_subtypes<-Main_Dataset[Main_Dataset$Status=="1",]
```

#####Overlook of the distribution of subtypes:
```{r,error=TRUE}
summary(Data_only_subtypes$Diagnosis)
```

####Partition the dataset containing only Congo-Red positive samples into a training dataset and a test dataset. The datasets are split randomly 70:30.
```{r,error=TRUE}
set.seed(1)
index_diagnosis<- createDataPartition(Data_only_subtypes$Diagnosis, p = 0.7, list = FALSE) #caret package must be loaded.
Training_Data_only_subtypes <- Data_only_subtypes[index_diagnosis, ]
Test_Data_only_subtypes <- Data_only_subtypes[-index_diagnosis, ]
```


#Feature Selection

###Boruta Feature Selection - Identification of Congo-Red positive samples:
```{r,error=TRUE}
set.seed(1)
Boruta_Output_Congo<-Boruta(Status~. ,data=Training_Data[c(2,4:ncol(Training_Data))],doTrace=0) #Boruta package must be loaded
Boruta_Attributes_Extraction_Congo <- getSelectedAttributes(Boruta_Output_Congo, withTentative = FALSE)
print(Boruta_Attributes_Extraction_Congo) 
Boruta_Congo_Statistics<-attStats(Boruta_Output_Congo) #
```

####Extract Top 10 most important proteins for differentiating between Congo-Red positive and negative samples:
```{r}
Training_Data_only_Top10<- Training_Data %>% select(Patients,Status, Diagnosis,P10909,P02743,P06727,P04004,P02649,P02748,P12109,P12110,P12111,P23142)
```

###Boruta Feature Selection - Identification of proteins important for subtype classification:
```{r,error=TRUE}
set.seed(1)
Boruta_Output_Subtype<-Boruta(Diagnosis~. ,data=Training_Data_only_subtypes[c(3,4:ncol(Training_Data_only_subtypes))],doTrace=0) #Boruta package must be loaded
Boruta_Attributes_Extraction_Subtype <- getSelectedAttributes(Boruta_Output_Subtype, withTentative = FALSE)
print(Boruta_Attributes_Extraction_Subtype) 
Boruta_Subtype_Statistics<-attStats(Boruta_Output_Subtype)
```

####Subset dataset to only include subtype-associated proteins
```{r,error=TRUE}
Training_Data_Subtypes<- Training_Data_only_subtypes %>% select(Patients,Status, Diagnosis,P02766,P0DOY2,P0DOX8,P0DOX7,P01834,P0DJI8,P0DJI9)
Test_Data_Subtypes<-Test_Data_only_subtypes %>% select(Patients,Status, Diagnosis,P02766,P0DOY2,P0DOX8,P0DOX7,P01834,P0DJI8,P0DJI9)
```


#Model Building

###Step 1

####APOA4
```{r,error=TRUE}
#Establish optimal model parameters
set.seed(1)
Model1_tuning_parameters_APOA4_only<-tune.svm(Status~., data = Training_Data_only_Top10[c(2,6)], sampling = "fix", gamma = 2^c(-15,-10,-5,5,10,15), cost = 2^c(-15,-10,-5,5,10,15))
Model1_tuning_parameters_APOA4_only$best.parameters

#Train classifier
set.seed(1)
SVM_model1_APOA4_only<-svm(Status~.,data=Training_Data_only_Top10[c(2,6)],method="C-classification",kernel="linear", gamma=Model1_tuning_parameters_APOA4_only$best.parameters$gamma, cost=Model1_tuning_parameters_APOA4_only$best.parameters$cost,decision.values=TRUE)

#Test classifier on Training Dataset
Train_model1_only_APOA4<-predict(SVM_model1_APOA4_only,newdata =Training_Data_only_Top10[4:13],decision.values=TRUE)
confusionMatrix(Train_model1_only_APOA4,Training_Data_only_Top10$Status,positive="1")

#Test classifier on Test Dataset
Test_model1_only_APOA4<-predict(SVM_model1_APOA4_only,newdata = Test_Data[4:ncol(Test_Data)],decision.values=TRUE)
confusionMatrix(Test_model1_only_APOA4,Test_Data$Status,positive="1")
```

####APOE
```{r,error=TRUE}

#Establish optimal model parameters
set.seed(1)
Model1_tuning_parameters_APOE_only<-tune.svm(Status~., data = Training_Data_only_Top10[c(2,8)], sampling = "fix", gamma = 2^c(-15,-10,-5,5,10,15), cost = 2^c(-15,-10,-5,5,10,15))
Model1_tuning_parameters_APOE_only$best.parameters

#Train classifier
set.seed(1)
SVM_model1_APOE_only<-svm(Status~.,data=Training_Data_only_Top10[c(2,8)],method="C-classification",kernel="linear", gamma=Model1_tuning_parameters_APOE_only$best.parameters$gamma, cost=Model1_tuning_parameters_APOE_only$best.parameters$cost,decision.values=TRUE)

#Test classifier on Training Dataset
Train_model1_only_APOE<-predict(SVM_model1_APOE_only,newdata =Training_Data_only_Top10[4:13],decision.values=TRUE)
confusionMatrix(Train_model1_only_APOE,Training_Data_only_Top10$Status,positive="1")

#Test classifier on Test Dataset
Test_model1_only_APOE<-predict(SVM_model1_APOE_only,newdata = Test_Data[4:ncol(Test_Data)],decision.values=TRUE)
confusionMatrix(Test_model1_only_APOE,Test_Data$Status,positive="1")
```

####SAP
```{r,error=TRUE}
#Establish optimal model parameters
set.seed(1)
Model1_tuning_parameters_SAP_only<-tune.svm(Status~., data = Training_Data_only_Top10[c(2,5)], sampling = "fix", gamma = 2^c(-15,-10,-5,5,10,15), cost = 2^c(-15,-10,-5,5,10,15))
Model1_tuning_parameters_SAP_only$best.parameters

#Train classifier
set.seed(1)
SVM_model1_SAP_only<-svm(Status~.,data=Training_Data_only_Top10[c(2,5)],method="C-classification",kernel="linear", gamma=Model1_tuning_parameters_SAP_only$best.parameters$gamma, cost=Model1_tuning_parameters_SAP_only$best.parameters$cost,decision.values=TRUE)

#Test classifier on Training Dataset
Train_model1_only_SAP<-predict(SVM_model1_SAP_only,newdata =Training_Data_only_Top10[4:13],decision.values=TRUE)
confusionMatrix(Train_model1_only_SAP,Training_Data_only_Top10$Status,positive="1")

#Test classifier on Test Dataset
Test_model1_only_SAP<-predict(SVM_model1_SAP_only,newdata = Test_Data[4:ncol(Test_Data)],decision.values=TRUE)
confusionMatrix(Test_model1_only_SAP,Test_Data$Status,positive="1")
```
####Clusterin
```{r,error=TRUE}
#Establish optimal model parameters
set.seed(1)
Model1_tuning_parameters_Clusterin_only<-tune.svm(Status~., data = Training_Data_only_Top10[c(2,4)], sampling = "fix", gamma = 2^c(-15,-10,-5,5,10,15), cost = 2^c(-15,-10,-5,5,10,15))
Model1_tuning_parameters_Clusterin_only$best.parameters

#Train classifier
set.seed(1)
SVM_model1_Clusterin_only<-svm(Status~.,data=Training_Data_only_Top10[c(2,4)],method="C-classification",kernel="linear", gamma=Model1_tuning_parameters_Clusterin_only$best.parameters$gamma, cost=Model1_tuning_parameters_Clusterin_only$best.parameters$cost,decision.values=TRUE)

#Test classifier on Training Dataset
Train_model1_only_Clusterin<-predict(SVM_model1_Clusterin_only,newdata =Training_Data_only_Top10[4:13],decision.values=TRUE)
confusionMatrix(Train_model1_only_Clusterin,Training_Data_only_Top10$Status,positive="1")

#Test classifier on Test Dataset
Test_model1_only_Clusterin<-predict(SVM_model1_Clusterin_only,newdata = Test_Data[4:ncol(Test_Data)],decision.values=TRUE)
confusionMatrix(Test_model1_only_Clusterin,Test_Data$Status,positive="1")
```
####Vitronectin
```{r,error=TRUE}
#Establish optimal model parameters
set.seed(1)
Model1_tuning_parameters_Vitronectin_only<-tune.svm(Status~., data = Training_Data_only_Top10[c(2,7)], sampling = "fix", gamma = 2^c(-15,-10,-5,5,10,15), cost = 2^c(-15,-10,-5,5,10,15))
Model1_tuning_parameters_Vitronectin_only$best.parameters

#Train classifier
set.seed(1)
SVM_model1_Vitronectin_only<-svm(Status~.,data=Training_Data_only_Top10[c(2,7)],method="C-classification",kernel="linear", gamma=Model1_tuning_parameters_Vitronectin_only$best.parameters$gamma, cost=Model1_tuning_parameters_Vitronectin_only$best.parameters$cost,decision.values=TRUE)

#Test classifier on Training Dataset
Train_model1_only_Vitronectin<-predict(SVM_model1_Vitronectin_only,newdata =Training_Data_only_Top10[4:13],decision.values=TRUE)
confusionMatrix(Train_model1_only_Vitronectin,Training_Data_only_Top10$Status,positive="1")

#Test classifier on Test Dataset
Test_model1_only_Vitronectin<-predict(SVM_model1_Vitronectin_only,newdata = Test_Data[4:ncol(Test_Data)],decision.values=TRUE)
confusionMatrix(Test_model1_only_Vitronectin,Test_Data$Status,positive="1")
```
####Complement component C9
```{r,error=TRUE}
#Establish optimal model parameters
set.seed(1)
Model1_tuning_parameters_Complement_C9_only<-tune.svm(Status~., data = Training_Data_only_Top10[c(2,9)], sampling = "fix", gamma = 2^c(-15,-10,-5,5,10,15), cost = 2^c(-15,-10,-5,5,10,15))
Model1_tuning_parameters_Complement_C9_only$best.parameters

#Train classifier
set.seed(1)
SVM_model1_Complement_C9_only<-svm(Status~.,data=Training_Data_only_Top10[c(2,9)],method="C-classification",kernel="linear", gamma=Model1_tuning_parameters_Complement_C9_only$best.parameters$gamma, cost=Model1_tuning_parameters_Complement_C9_only$best.parameters$cost,decision.values=TRUE)

#Test classifier on Training Dataset
Train_model1_only_Complement_C9<-predict(SVM_model1_Complement_C9_only,newdata =Training_Data_only_Top10[4:13],decision.values=TRUE)
confusionMatrix(Train_model1_only_Complement_C9,Training_Data_only_Top10$Status,positive="1")

#Test classifier on Test Dataset
Test_model1_only_Complement_C9<-predict(SVM_model1_Complement_C9_only,newdata = Test_Data[4:ncol(Test_Data)],decision.values=TRUE)
confusionMatrix(Test_model1_only_Complement_C9,Test_Data$Status,positive="1")
```
####COL6A1
```{r,error=TRUE}
#Establish optimal model parameters
set.seed(1)
Model1_tuning_parameters_COL6A1_only<-tune.svm(Status~., data = Training_Data_only_Top10[c(2,10)], sampling = "fix", gamma = 2^c(-15,-10,-5,5,10,15), cost = 2^c(-15,-10,-5,5,10,15))
Model1_tuning_parameters_COL6A1_only$best.parameters

#Train classifier
set.seed(1)
SVM_model1_COL6A1_only<-svm(Status~.,data=Training_Data_only_Top10[c(2,10)],method="C-classification",kernel="linear", gamma=Model1_tuning_parameters_COL6A1_only$best.parameters$gamma, cost=Model1_tuning_parameters_COL6A1_only$best.parameters$cost,decision.values=TRUE)

#Test classifier on Training Dataset
Train_model1_only_COL6A1<-predict(SVM_model1_COL6A1_only,newdata =Training_Data_only_Top10[4:13],decision.values=TRUE)
confusionMatrix(Train_model1_only_COL6A1,Training_Data_only_Top10$Status,positive="1")

#Test classifier on Test Dataset
Test_model1_only_COL6A1<-predict(SVM_model1_COL6A1_only,newdata = Test_Data[4:ncol(Test_Data)],decision.values=TRUE)
confusionMatrix(Test_model1_only_COL6A1,Test_Data$Status,positive="1")
```
####COL6A2
```{r,error=TRUE}
#Establish optimal model parameters
set.seed(1)
Model1_tuning_parameters_COL6A2_only<-tune.svm(Status~., data = Training_Data_only_Top10[c(2,11)], sampling = "fix", gamma = 2^c(-15,-10,-5,5,10,15), cost = 2^c(-15,-10,-5,5,10,15))
Model1_tuning_parameters_COL6A2_only$best.parameters

#Train classifier
set.seed(1)
SVM_model1_COL6A2_only<-svm(Status~.,data=Training_Data_only_Top10[c(2,11)],method="C-classification",kernel="linear", gamma=Model1_tuning_parameters_COL6A2_only$best.parameters$gamma, cost=Model1_tuning_parameters_COL6A2_only$best.parameters$cost,decision.values=TRUE)

#Test classifier on Training Dataset
Train_model1_only_COL6A2<-predict(SVM_model1_COL6A2_only,newdata =Training_Data_only_Top10[4:13],decision.values=TRUE)
confusionMatrix(Train_model1_only_COL6A2,Training_Data_only_Top10$Status,positive="1")

#Test classifier on Test Dataset
Test_model1_only_COL6A2<-predict(SVM_model1_COL6A2_only,newdata = Test_Data[4:ncol(Test_Data)],decision.values=TRUE)
confusionMatrix(Test_model1_only_COL6A2,Test_Data$Status,positive="1")
```
####COL6A3
```{r,error=TRUE}
#Establish optimal model parameters
set.seed(1)
Model1_tuning_parameters_COL6A3_only<-tune.svm(Status~., data = Training_Data_only_Top10[c(2,12)], sampling = "fix", gamma = 2^c(-15,-10,-5,5,10,15), cost = 2^c(-15,-10,-5,5,10,15))
Model1_tuning_parameters_COL6A3_only$best.parameters

#Train classifier
set.seed(1)
SVM_model1_COL6A3_only<-svm(Status~.,data=Training_Data_only_Top10[c(2,12)],method="C-classification",kernel="linear", gamma=Model1_tuning_parameters_COL6A3_only$best.parameters$gamma, cost=Model1_tuning_parameters_COL6A3_only$best.parameters$cost,decision.values=TRUE)

#Test classifier on Training Dataset
Train_model1_only_COL6A3<-predict(SVM_model1_COL6A3_only,newdata =Training_Data_only_Top10[4:13],decision.values=TRUE)
confusionMatrix(Train_model1_only_COL6A3,Training_Data_only_Top10$Status,positive="1")

#Test classifier on Test Dataset
Test_model1_only_COL6A3<-predict(SVM_model1_COL6A3_only,newdata = Test_Data[4:ncol(Test_Data)],decision.values=TRUE)
confusionMatrix(Test_model1_only_COL6A3,Test_Data$Status,positive="1")
```
####Fibulin-1
```{r,error=TRUE}
#Establish optimal model parameters
set.seed(1)
Model1_tuning_parameters_Fibulin1_only<-tune.svm(Status~., data = Training_Data_only_Top10[c(2,13)], sampling = "fix", gamma = 2^c(-15,-10,-5,5,10,15), cost = 2^c(-15,-10,-5,5,10,15))
Model1_tuning_parameters_Fibulin1_only$best.parameters

#Train classifier
set.seed(1)
SVM_model1_Fibulin1_only<-svm(Status~.,data=Training_Data_only_Top10[c(2,13)],method="C-classification",kernel="linear", gamma=Model1_tuning_parameters_Fibulin1_only$best.parameters$gamma, cost=Model1_tuning_parameters_Fibulin1_only$best.parameters$cost,decision.values=TRUE)

#Test classifier on Training Dataset
Train_model1_only_Fibulin1<-predict(SVM_model1_Fibulin1_only,newdata =Training_Data_only_Top10[4:13],decision.values=TRUE)
confusionMatrix(Train_model1_only_Fibulin1,Training_Data_only_Top10$Status,positive="1")

#Test classifier on Test Dataset
Test_model1_only_Fibulin1<-predict(SVM_model1_Fibulin1_only,newdata = Test_Data[4:ncol(Test_Data)],decision.values=TRUE)
confusionMatrix(Test_model1_only_Fibulin1,Test_Data$Status,positive="1")
```
####APOA4 and APOE

```{r,error=TRUE}
#Establish optimal model parameters
set.seed(1)
Model1_tuning_parameters_APOA4_and_APOE_only<-tune.svm(Status~., data = Training_Data_only_Top10[c(2,6,8)], sampling = "fix", gamma = 2^c(-15,-10,-5,5,10,15), cost = 2^c(-15,-10,-5,5,10,15))
Model1_tuning_parameters_APOA4_and_APOE_only$best.parameters

#Train classifier
set.seed(1)
SVM_model1_APOA4_and_APOE_only<-svm(Status~.,data=Training_Data_only_Top10[c(2,6,8)],method="C-classification",kernel="linear", gamma=Model1_tuning_parameters_APOA4_and_APOE_only$best.parameters$gamma, cost=Model1_tuning_parameters_APOA4_and_APOE_only$best.parameters$cost,decision.values=TRUE)

#Test classifier on Training Dataset
Train_model1_only_APOA4_and_APOE<-predict(SVM_model1_APOA4_and_APOE_only,newdata =Training_Data_only_Top10[4:13],decision.values=TRUE)
confusionMatrix(Train_model1_only_APOA4_and_APOE,Training_Data_only_Top10$Status,positive="1")

#Test classifier on Test Dataset
Test_model1_only_APOA4_and_APOE<-predict(SVM_model1_APOA4_and_APOE_only,newdata = Test_Data[4:ncol(Test_Data)],decision.values=TRUE)
confusionMatrix(Test_model1_only_APOA4_and_APOE,Test_Data$Status,positive="1")

```
#####APOA4, APOE and Clusterin
```{r,error=TRUE}
#Establish optimal model parameters
set.seed(1)
Model1_tuning_parameters_APOA4_APOE_Clusterin_only<-tune.svm(Status~., data = Training_Data_only_Top10[c(2,6,8,4)], sampling = "fix", gamma = 2^c(-15,-10,-5,5,10,15), cost = 2^c(-15,-10,-5,5,10,15))
Model1_tuning_parameters_APOA4_APOE_Clusterin_only$best.parameters

#Train classifier
set.seed(1)
SVM_model1_APOA4_APOE_Clusterin_only<-svm(Status~.,data=Training_Data_only_Top10[c(2,6,8,4)],method="C-classification",kernel="linear", gamma=Model1_tuning_parameters_APOA4_APOE_Clusterin_only$best.parameters$gamma, cost=Model1_tuning_parameters_APOA4_APOE_Clusterin_only$best.parameters$cost,decision.values=TRUE)

#Test classifier on Training Dataset
Train_model1_only_APOA4_APOE_Clusterin<-predict(SVM_model1_APOA4_APOE_Clusterin_only,newdata =Training_Data_only_Top10[4:13],decision.values=TRUE)
confusionMatrix(Train_model1_only_APOA4_APOE_Clusterin,Training_Data_only_Top10$Status,positive="1")

#Test classifier on Test Dataset
Test_model1_only_APOA4_APOE_Clusterin<-predict(SVM_model1_APOA4_APOE_Clusterin_only,newdata = Test_Data[4:ncol(Test_Data)],decision.values=TRUE)
confusionMatrix(Test_model1_only_APOA4_APOE_Clusterin,Test_Data$Status,positive="1")
```
#####APOA4, APOE and Vitronectin
```{r,error=TRUE}
#Establish optimal model parameters
set.seed(1)
Model1_tuning_parameters_APOA4_APOE_Vitronectin_only<-tune.svm(Status~., data = Training_Data_only_Top10[c(2,6,8,7)], sampling = "fix", gamma = 2^c(-15,-10,-5,5,10,15), cost = 2^c(-15,-10,-5,5,10,15))
Model1_tuning_parameters_APOA4_APOE_Vitronectin_only$best.parameters

#Train classifier
set.seed(1)
SVM_model1_APOA4_APOE_Vitronectin_only<-svm(Status~.,data=Training_Data_only_Top10[c(2,6,8,7)],method="C-classification",kernel="linear", gamma=Model1_tuning_parameters_APOA4_APOE_Vitronectin_only$best.parameters$gamma, cost=Model1_tuning_parameters_APOA4_APOE_Vitronectin_only$best.parameters$cost,decision.values=TRUE)

#Test classifier on Training Dataset
Train_model1_only_APOA4_APOE_Vitronectin<-predict(SVM_model1_APOA4_APOE_Vitronectin_only,newdata =Training_Data_only_Top10[4:13],decision.values=TRUE)
confusionMatrix(Train_model1_only_APOA4_APOE_Vitronectin,Training_Data_only_Top10$Status,positive="1")

#Test classifier on Test Dataset
Test_model1_only_APOA4_APOE_Vitronectin<-predict(SVM_model1_APOA4_APOE_Vitronectin_only,newdata = Test_Data[4:ncol(Test_Data)],decision.values=TRUE)
confusionMatrix(Test_model1_only_APOA4_APOE_Vitronectin,Test_Data$Status,positive="1")
```
#####APOA4, APOE and Complement component C9
```{r,error=TRUE}
#Establish optimal model parameters
set.seed(1)
Model1_tuning_parameters_APOA4_APOE_C9_only<-tune.svm(Status~., data = Training_Data_only_Top10[c(2,6,8,9)], sampling = "fix", gamma = 2^c(-15,-10,-5,5,10,15), cost = 2^c(-15,-10,-5,5,10,15))
Model1_tuning_parameters_APOA4_APOE_C9_only$best.parameters

#Train classifier
set.seed(1)
SVM_model1_APOA4_APOE_C9_only<-svm(Status~.,data=Training_Data_only_Top10[c(2,6,8,9)],method="C-classification",kernel="linear", gamma=Model1_tuning_parameters_APOA4_APOE_C9_only$best.parameters$gamma, cost=Model1_tuning_parameters_APOA4_APOE_C9_only$best.parameters$cost,decision.values=TRUE)

#Test classifier on Training Dataset
Train_model1_only_APOA4_APOE_C9<-predict(SVM_model1_APOA4_APOE_C9_only,newdata =Training_Data_only_Top10[4:13],decision.values=TRUE)
confusionMatrix(Train_model1_only_APOA4_APOE_C9,Training_Data_only_Top10$Status,positive="1")

#Test classifier on Test Dataset
Test_model1_only_APOA4_APOE_C9<-predict(SVM_model1_APOA4_APOE_C9_only,newdata = Test_Data[4:ncol(Test_Data)],decision.values=TRUE)
confusionMatrix(Test_model1_only_APOA4_APOE_C9,Test_Data$Status,positive="1")
```
#####APOA4, APOE and COL6A1
```{r,error=TRUE}
#Establish optimal model parameters
set.seed(1)
Model1_tuning_parameters_APOA4_APOE_COL6A1_only<-tune.svm(Status~., data = Training_Data_only_Top10[c(2,6,8,10)], sampling = "fix", gamma = 2^c(-15,-10,-5,5,10,15), cost = 2^c(-15,-10,-5,5,10,15))
Model1_tuning_parameters_APOA4_APOE_COL6A1_only$best.parameters

#Train classifier
set.seed(1)
SVM_model1_APOA4_APOE_COL6A1_only<-svm(Status~.,data=Training_Data_only_Top10[c(2,6,8,10)],method="C-classification",kernel="linear", gamma=Model1_tuning_parameters_APOA4_APOE_COL6A1_only$best.parameters$gamma, cost=Model1_tuning_parameters_APOA4_APOE_COL6A1_only$best.parameters$cost,decision.values=TRUE)

#Test classifier on Training Dataset
Train_model1_only_APOA4_APOE_COL6A1<-predict(SVM_model1_APOA4_APOE_COL6A1_only,newdata =Training_Data_only_Top10[4:13],decision.values=TRUE)
confusionMatrix(Train_model1_only_APOA4_APOE_COL6A1,Training_Data_only_Top10$Status,positive="1")

#Test classifier on Test Dataset
Test_model1_only_APOA4_APOE_COL6A1<-predict(SVM_model1_APOA4_APOE_COL6A1_only,newdata = Test_Data[4:ncol(Test_Data)],decision.values=TRUE)
confusionMatrix(Test_model1_only_APOA4_APOE_COL6A1,Test_Data$Status,positive="1")
```
#####APOA4, APOE and COL6A2
```{r,error=TRUE}
#Establish optimal model parameters
set.seed(1)
Model1_tuning_parameters_APOA4_APOE_COL6A2_only<-tune.svm(Status~., data = Training_Data_only_Top10[c(2,6,8,11)], sampling = "fix", gamma = 2^c(-15,-10,-5,5,10,15), cost = 2^c(-15,-10,-5,5,10,15))
Model1_tuning_parameters_APOA4_APOE_COL6A2_only$best.parameters

#Train classifier
set.seed(1)
SVM_model1_APOA4_APOE_COL6A2_only<-svm(Status~.,data=Training_Data_only_Top10[c(2,6,8,11)],method="C-classification",kernel="linear", gamma=Model1_tuning_parameters_APOA4_APOE_COL6A2_only$best.parameters$gamma, cost=Model1_tuning_parameters_APOA4_APOE_COL6A2_only$best.parameters$cost,decision.values=TRUE)

#Test classifier on Training Dataset
Train_model1_only_APOA4_APOE_COL6A2<-predict(SVM_model1_APOA4_APOE_COL6A2_only,newdata =Training_Data_only_Top10[4:13],decision.values=TRUE)
confusionMatrix(Train_model1_only_APOA4_APOE_COL6A2,Training_Data_only_Top10$Status,positive="1")

#Test classifier on Test Dataset
Test_model1_only_APOA4_APOE_COL6A2<-predict(SVM_model1_APOA4_APOE_COL6A2_only,newdata = Test_Data[4:ncol(Test_Data)],decision.values=TRUE)
confusionMatrix(Test_model1_only_APOA4_APOE_COL6A2,Test_Data$Status,positive="1")
```
#####APOA4, APOE and COL6A3
```{r,error=TRUE}
#Establish optimal model parameters
set.seed(1)
Model1_tuning_parameters_APOA4_APOE_COL6A3_only<-tune.svm(Status~., data = Training_Data_only_Top10[c(2,6,8,12)], sampling = "fix", gamma = 2^c(-15,-10,-5,5,10,15), cost = 2^c(-15,-10,-5,5,10,15))
Model1_tuning_parameters_APOA4_APOE_COL6A3_only$best.parameters

#Train classifier
set.seed(1)
SVM_model1_APOA4_APOE_COL6A3_only<-svm(Status~.,data=Training_Data_only_Top10[c(2,6,8,12)],method="C-classification",kernel="linear", gamma=Model1_tuning_parameters_APOA4_APOE_COL6A3_only$best.parameters$gamma, cost=Model1_tuning_parameters_APOA4_APOE_COL6A3_only$best.parameters$cost,decision.values=TRUE)

#Test classifier on Training Dataset
Train_model1_only_APOA4_APOE_COL6A3<-predict(SVM_model1_APOA4_APOE_COL6A3_only,newdata =Training_Data_only_Top10[4:13],decision.values=TRUE)
confusionMatrix(Train_model1_only_APOA4_APOE_COL6A3,Training_Data_only_Top10$Status,positive="1")

#Test classifier on Test Dataset
Test_model1_only_APOA4_APOE_COL6A3<-predict(SVM_model1_APOA4_APOE_COL6A3_only,newdata = Test_Data[4:ncol(Test_Data)],decision.values=TRUE)
confusionMatrix(Test_model1_only_APOA4_APOE_COL6A3,Test_Data$Status,positive="1")
```
##### APOA4, APOE and Fibulin-1
```{r,error=TRUE}
#Establish optimal model parameters
set.seed(1)
Model1_tuning_parameters_APOA4_APOE_Fibulin1_only<-tune.svm(Status~., data = Training_Data_only_Top10[c(2,6,8,13)], sampling = "fix", gamma = 2^c(-15,-10,-5,5,10,15), cost = 2^c(-15,-10,-5,5,10,15))
Model1_tuning_parameters_APOA4_APOE_Fibulin1_only$best.parameters

#Train classifier
set.seed(1)
SVM_model1_APOA4_APOE_Fibulin1_only<-svm(Status~.,data=Training_Data_only_Top10[c(2,6,8,13)],method="C-classification",kernel="linear", gamma=Model1_tuning_parameters_APOA4_APOE_Fibulin1_only$best.parameters$gamma, cost=Model1_tuning_parameters_APOA4_APOE_Fibulin1_only$best.parameters$cost,decision.values=TRUE)

#Test classifier on Training Dataset
Train_model1_only_APOA4_APOE_Fibulin1<-predict(SVM_model1_APOA4_APOE_Fibulin1_only,newdata =Training_Data_only_Top10[4:13],decision.values=TRUE)
confusionMatrix(Train_model1_only_APOA4_APOE_Fibulin1,Training_Data_only_Top10$Status,positive="1")

#Test classifier on Test Dataset
Test_model1_only_APOA4_APOE_Fibulin1<-predict(SVM_model1_APOA4_APOE_Fibulin1_only,newdata = Test_Data[4:ncol(Test_Data)],decision.values=TRUE)
confusionMatrix(Test_model1_only_APOA4_APOE_Fibulin1,Test_Data$Status,positive="1")
```
####APOA4, APOE and SAP
```{r,error=TRUE}
#Establish optimal model parameters
set.seed(1)
Model1_tuning_parameters_APOA4_APOE_SAP_only<-tune.svm(Status~., data = Training_Data_only_Top10[c(2,6,8,5)], sampling = "fix", gamma = 2^c(-15,-10,-5,5,10,15), cost = 2^c(-15,-10,-5,5,10,15))
Model1_tuning_parameters_APOA4_APOE_SAP_only$best.parameters

#Train classifier
set.seed(1)
SVM_model1_APOA4_APOE_SAP_only<-svm(Status~.,data=Training_Data_only_Top10[c(2,6,8,5)],method="C-classification",kernel="linear", gamma=Model1_tuning_parameters_APOA4_APOE_SAP_only$best.parameters$gamma, cost=Model1_tuning_parameters_APOA4_APOE_SAP_only$best.parameters$cost,decision.values=TRUE)

#Test classifier on Training Dataset
Train_model1_only_APOA4_APOE_SAP<-predict(SVM_model1_APOA4_APOE_SAP_only,newdata =Training_Data_only_Top10[4:13],decision.values=TRUE)
confusionMatrix(Train_model1_only_APOA4_APOE_SAP,Training_Data_only_Top10$Status,positive="1")

#Test classifier on Test Dataset
Test_model1_only_APOA4_APOE_SAP<-predict(SVM_model1_APOA4_APOE_SAP_only,newdata = Test_Data[4:ncol(Test_Data)],decision.values=TRUE)
confusionMatrix(Test_model1_only_APOA4_APOE_SAP,Test_Data$Status,positive="1")
```


####Validate results on the blinded, independent validation dataset:

#####APOA4
```{r}
Validation_of_classifier_w_only_APOA4<-predict(SVM_model1_APOA4_only,newdata = Validation_Dataset[4:ncol(Validation_Dataset)],decision.values=TRUE)
summary(Validation_of_classifier_w_only_APOA4)
```

#####APOE
```{r}
Validation_of_classifier_w_only_APOE<-predict(SVM_model1_APOE_only,newdata = Validation_Dataset[4:ncol(Validation_Dataset)],decision.values=TRUE)
summary(Validation_of_classifier_w_only_APOE)
```
#####SAP
```{r}
Validation_of_classifier_w_only_SAP<-predict(SVM_model1_SAP_only,newdata = Validation_Dataset[4:ncol(Validation_Dataset)],decision.values=TRUE)
summary(Validation_of_classifier_w_only_SAP)
```
#####Clusterin
```{r}
Validation_of_classifier_w_only_Clusterin<-predict(SVM_model1_Clusterin_only,newdata = Validation_Dataset[4:ncol(Validation_Dataset)],decision.values=TRUE)
summary(Validation_of_classifier_w_only_Clusterin)
```
#####Vitronectin
```{r}
Validation_of_classifier_w_only_Vitronectin<-predict(SVM_model1_Vitronectin_only,newdata = Validation_Dataset[4:ncol(Validation_Dataset)],decision.values=TRUE)
summary(Validation_of_classifier_w_only_Vitronectin)
```
#####Complement component C9
```{r}
Validation_of_classifier_w_only_C9<-predict(SVM_model1_Complement_C9_only,newdata = Validation_Dataset[4:ncol(Validation_Dataset)],decision.values=TRUE)
summary(Validation_of_classifier_w_only_C9)
```
#####COL6A1
```{r}
Validation_of_classifier_w_only_COL6A1<-predict(SVM_model1_COL6A1_only,newdata = Validation_Dataset[4:ncol(Validation_Dataset)],decision.values=TRUE)
summary(Validation_of_classifier_w_only_COL6A1)
```
#####COL6A2
```{r}
Validation_of_classifier_w_only_COL6A2<-predict(SVM_model1_COL6A2_only,newdata = Validation_Dataset[4:ncol(Validation_Dataset)],decision.values=TRUE)
summary(Validation_of_classifier_w_only_COL6A2)
```
#####COL6A3
```{r}
Validation_of_classifier_w_only_COL6A3<-predict(SVM_model1_COL6A3_only,newdata = Validation_Dataset[4:ncol(Validation_Dataset)],decision.values=TRUE)
summary(Validation_of_classifier_w_only_COL6A3)
```
#####Fibulin-1
```{r}
Validation_of_classifier_w_only_Fibulin1<-predict(SVM_model1_Fibulin1_only,newdata = Validation_Dataset[4:ncol(Validation_Dataset)],decision.values=TRUE)
summary(Validation_of_classifier_w_only_Fibulin1)
```
#####APOA4 and APOE
```{r}
Validation_of_classifier_w_only_APOA4_and_APOE<-predict(SVM_model1_APOA4_and_APOE_only,newdata = Validation_Dataset[4:ncol(Validation_Dataset)],decision.values=TRUE)
summary(Validation_of_classifier_w_only_APOA4_and_APOE)
```
######APOA4, APOE and Clusterin
```{r}
Validation_of_classifier_w_only_APOA4_APOE_Clusterin<-predict(SVM_model1_APOA4_APOE_Clusterin_only,newdata = Validation_Dataset[4:ncol(Validation_Dataset)],decision.values=TRUE)
summary(Validation_of_classifier_w_only_APOA4_APOE_Clusterin)
```
######APOA4, APOE and Vitronectin
```{r}
Validation_of_classifier_w_only_APOA4_APOE_Vitronectin<-predict(SVM_model1_APOA4_APOE_Vitronectin_only,newdata = Validation_Dataset[4:ncol(Validation_Dataset)],decision.values=TRUE)
summary(Validation_of_classifier_w_only_APOA4_APOE_Vitronectin)
```
######APOA4, APOE and Complement component C9
```{r}
Validation_of_classifier_w_only_APOA4_APOE_C9<-predict(SVM_model1_APOA4_APOE_C9_only,newdata = Validation_Dataset[4:ncol(Validation_Dataset)],decision.values=TRUE)
summary(Validation_of_classifier_w_only_APOA4_APOE_C9)
```
######APOA4, APOE and COL6A1
```{r}
Validation_of_classifier_w_only_APOA4_APOE_COL6A1<-predict(SVM_model1_APOA4_APOE_COL6A1_only,newdata = Validation_Dataset[4:ncol(Validation_Dataset)],decision.values=TRUE)
summary(Validation_of_classifier_w_only_APOA4_APOE_COL6A1)
```
######APOA4, APOE and COL6A2
```{r}
Validation_of_classifier_w_only_APOA4_APOE_COL6A2<-predict(SVM_model1_APOA4_APOE_COL6A2_only,newdata = Validation_Dataset[4:ncol(Validation_Dataset)],decision.values=TRUE)
summary(Validation_of_classifier_w_only_APOA4_APOE_COL6A2)
```
######APOA4, APOE and COL6A3
```{r}
Validation_of_classifier_w_only_APOA4_APOE_COL6A3<-predict(SVM_model1_APOA4_APOE_COL6A3_only,newdata = Validation_Dataset[4:ncol(Validation_Dataset)],decision.values=TRUE)
summary(Validation_of_classifier_w_only_APOA4_APOE_COL6A3)
```
######APOA4, APOE and Fibulin-1
```{r}
Validation_of_classifier_w_only_APOA4_APOE_Fibulin1<-predict(SVM_model1_APOA4_APOE_Fibulin1_only,newdata = Validation_Dataset[4:ncol(Validation_Dataset)],decision.values=TRUE)
summary(Validation_of_classifier_w_only_APOA4_APOE_Fibulin1)
```

#####APOA4, APOE and SAP
```{r}
Validation_of_classifier_w_only_APOA4_APOE_SAP<-predict(SVM_model1_APOA4_APOE_SAP_only,newdata = Validation_Dataset[4:ncol(Validation_Dataset)],decision.values=TRUE)
summary(Validation_of_classifier_w_only_APOA4_APOE_SAP)
```



###Step 2
#### Model with all Boruta selected proteins
```{r,error=TRUE}
#Establish optimal model parameters
set.seed(1)
Model2_tuning_parameters<-tune.svm(Diagnosis~ ., data=Training_Data_Subtypes[3:10],gamma = 2^c(-15,-10,-5,0,5,10,15), cost = 2^c(-15,-10,-5,0,5,10,15), sampling = "fix")
Model2_tuning_parameters$best.parameters

#Train classifier
SVM_Subtype_Classifier<-svm(Diagnosis~.,data=Training_Data_Subtypes[3:10],method="C-classification",kernel="linear",gamma=Model2_tuning_parameters$best.parameters$gamma	,cost=Model2_tuning_parameters$best.parameters$cost ,decision.values=TRUE)

#Test classifier on Training Dataset
set.seed(1)
summary(SVM_Subtype_Classifier)
Training_subtype_prediction<-predict(SVM_Subtype_Classifier,newdata = Training_Data_Subtypes[4:10],decision.values=TRUE)
confusionMatrix(Training_subtype_prediction,Training_Data_Subtypes$Diagnosis)

#Test classifier on Test Dataset
Test_subtype_prediction<-predict(SVM_Subtype_Classifier,newdata = Test_Data_Subtypes[4:10],decision.values=TRUE)
confusionMatrix(Test_subtype_prediction,Test_Data_Subtypes$Diagnosis)

#Validate classifier on Validation Dataset
Validation_subtype_prediction<-predict(SVM_Subtype_Classifier,newdata = Validation_Dataset[4:ncol(Validation_Dataset)],decision.values=TRUE)
confusionMatrix(Validation_subtype_prediction,Validation_Dataset$Diagnosis)

```

#### Model with only the known amyloid-associated proteins included
```{r,error=TRUE}
#Establish optimal model parameters
set.seed(1)
Model2_tuning_parameters1<-tune.svm(Diagnosis~ ., data=Training_Data_Subtypes[c(3,4,6,7,9)],gamma = 2^c(-15,-10,-5,0,5,10,15), cost = 2^c(-15,-10,-5,0,5,10,15), sampling = "fix")
Model2_tuning_parameters1$best.parameters

#Train classifier
SVM_Subtype_Classifier1<-svm(Diagnosis~.,data=Training_Data_Subtypes[c(3,4,6,7,9)],method="C-classification",kernel="linear",gamma=Model2_tuning_parameters1$best.parameters$gamma	,cost=Model2_tuning_parameters1$best.parameters$cost ,decision.values=TRUE)
summary(SVM_Subtype_Classifier1)
#Test classifier on Training Dataset
set.seed(1)
Training_subtype_prediction1<-predict(SVM_Subtype_Classifier1,newdata = Training_Data_Subtypes[4:10],decision.values=TRUE)
confusionMatrix(Training_subtype_prediction1,Training_Data_Subtypes$Diagnosis)

#Test classifier on Test Dataset
Test_subtype_prediction1<-predict(SVM_Subtype_Classifier1,newdata = Test_Data_Subtypes[4:10],decision.values=TRUE)
confusionMatrix(Test_subtype_prediction1,Test_Data_Subtypes$Diagnosis)

#Validate classifier on Validation Dataset
Validation_subtype_prediction1<-predict(SVM_Subtype_Classifier1,newdata = Validation_Dataset[4:ncol(Validation_Dataset)],decision.values=TRUE)
confusionMatrix(Validation_subtype_prediction1,Validation_Dataset$Diagnosis)

```
